---
title: "R_codes_all"
author: "HBZ"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#setwd("/Users/tmslab/Desktop/Research/CPM+craving+MUD/github_clone/CPM_EEG_craving/")
```

## load packages and data
```{r}
setwd("/Users/tmslab/Desktop/Research/CPM+craving+MUD/github_clone/CPM_EEG_craving/")
# Load necessary libraries
library(readxl)
library(dplyr)
library(knitr)
library(kableExtra)

# Load the data from the Excel file

data_train <- read_excel('data/all_data_2024.xlsx', sheet = "Training")
data_test <- read_excel("data/all_data_2024.xlsx", sheet = "Test")
```


## Correlation of different variables 
```{r}
library(Hmisc)

# Define the columns to include in the correlation matrix
columns_to_include <- c("age","edu_years","addiction_years", "withdraw_day",    "dose_month","craving_score","DSM","PSQI","BDI","BIS_total","BAI")
# Ensure these columns exist in the dataset
columns_to_include <- columns_to_include[columns_to_include %in% colnames(data_train)]

# Select the columns to perform the correlation analysis
data_to_correlate <- data_train %>% select(all_of(columns_to_include))

# Perform the Pearson correlation analysis
correlation_results <- rcorr(as.matrix(data_to_correlate), type = "pearson")

# Extract the correlation coefficients (r values) and round to 3 decimal places
correlation_matrix <- round(correlation_results$r, 2)

# Extract the p-values and round to 3 decimal places
p_value_matrix <- round(correlation_results$P, 3)

# Print the correlation coefficients
print("Correlation Coefficients with craving (r values):")
print(correlation_matrix["craving_score",])

# Print the p-values
print("P-values with craving:")
print(p_value_matrix["craving_score",])

```


## Comparison between training set and Test set profile
```{r}
# Add a column to distinguish between training and test sets
data_train$Set <- "Training"
data_test$Set <- "Test"

# Combine the data
combined_data <- bind_rows(data_train, data_test)

# Function to calculate mean, SD, t-value, and p-value
calculate_statistics <- function(data, variable) {
  summary_stats <- data %>%
    group_by(Set) %>%
    summarise(
      mean = mean(get(variable), na.rm = TRUE),
      sd = sd(get(variable), na.rm = TRUE)
    )
  
  t_test <- t.test(data[[variable]] ~ data$Set, var.equal = FALSE)
  t_value <- t_test$statistic
  p_value <- t_test$p.value
  
  result <- data.frame(
    Variable = variable,
    Training = paste0(format(round(summary_stats$mean[summary_stats$Set == "Training"], 2), nsmall = 2), " (", format(round(summary_stats$sd[summary_stats$Set == "Training"], 2), nsmall = 2), ")"),
    Test = paste0(format(round(summary_stats$mean[summary_stats$Set == "Test"], 2), nsmall = 2), " (", format(round(summary_stats$sd[summary_stats$Set == "Test"], 2), nsmall = 2), ")"),
    t_value = format(round(t_value, 2), nsmall = 2),
    p_value = format(round(p_value, 3), nsmall = 3)
  )
  
  return(result)
}

# List of variables to summarize
variables <- c("age", "edu_years", "addiction_years", "dose_month", "withdraw_day", "DSM", "PSQI", "BDI", "BIS_total", "BAI", "craving_score")

# Calculate statistics for each variable and combine into a single data frame
summary_table <- do.call(rbind, lapply(variables, calculate_statistics, data = combined_data))

# Create the table
kable(summary_table, format = "html", col.names = c("Variable", "Training (n=144)", "Test (n=47)", "t-value", "p-value")) %>%
  kable_styling(full_width = FALSE, position = "left")
```

## for other CPM related anlaysis, look at the Matlab folder and run in Matlab



## mediation analysis

### Data Prepartion

``` {r}
library(psych)
library(mediation)
library(diagram)
library(readxl)
library(DiagrammeR)
setwd("/Users/tmslab/Desktop/Research/CPM+craving+MUD/github_clone/CPM_EEG_craving/scripts/mediation_test")
source("process.R")

# Import the dataset
d <- read_excel("Mediation_test.xlsx", sheet = 'Combine')
d$BDI <- as.numeric(d$BDI)
```

## Data normalization
```{R}
#d <- d[1:144,]
# data visualization
hist(d$craving)
describe(d$craving)
hist(d$abstinence)
describe(d$abstinence)
hist(d$brain)
describe(d$brain)
hist(d$PSQI)
describe(d$PSQI)

# data normalization
d$craving_z <-  as.numeric(scale(d$craving, center = TRUE, scale = TRUE))
d$abstinence_z <-  as.numeric(scale(d$abstinence, center = TRUE, scale = TRUE))
d$brain_z <-  as.numeric(scale(d$brain, center = TRUE, scale = TRUE))
d$PSQI_z <-  as.numeric(scale(d$PSQI, center = TRUE, scale = TRUE))

d$BIS_z <-  as.numeric(scale(d$BIS_total, center = TRUE, scale = TRUE))
d$BISnoplan_z <-  as.numeric(scale(d$BIS_noplan, center = TRUE, scale = TRUE))
d$BISmotor_z <-  as.numeric(scale(d$BIS_motor, center = TRUE, scale = TRUE))
d$BISattention_z <-  as.numeric(scale(d$BIS_attention, center = TRUE, scale = TRUE))


d$BDI_z <-  as.numeric(scale(d$BDI, center = TRUE, scale = TRUE))
d$BAI_z <-  as.numeric(scale(d$BAI, center = TRUE, scale = TRUE))
```

## the relationship between Brain - BIS - craving
### parrelell mediation model
#### using BIS-total, BAI, and BDI
```{R}
# define X, Y, and M
d$X = d$brain_z
d$Y = d$craving_z

#Computing linear regression for PATH A1
Reg2PathA1 <- lm(BDI_z ~ X, data = d)
summary(Reg2PathA1)

#Computing linear regression for PATH A2
Reg2PathA2 <- lm(BIS_z ~ X, data = d)
summary(Reg2PathA2)

#Computing linear regression for PATH A3
Reg2PathA3 <- lm(BAI_z ~ X, data = d)
summary(Reg2PathA3)


#Conducting linear regression for PATH B1, B2, and C'
Reg2DirectEffect <- lm(Y ~ BDI_z + BIS_z + BAI_z + X, data = d)
summary(Reg2DirectEffect)

# Using PROCESS macro 

process(d, y="Y", x="X", m= c("BDI_z", "BIS_z","BAI_z"), model = 4,  effsize = 1,total = 1,  normal = 1, boot = 5000, seed = 9999)  

```

### abstinence -> brain  -> craving (brain only)

```{R}
# define X, Y, and M
d$X = d$abstinence_z
d$Y = d$craving_z

#Computing linear regression for PATH A
Reg2PathA <- lm(brain_z ~ X, data = d)
summary(Reg2PathA)

#Conducting linear regression for PATH B and C'
Reg2DirectEffect <- lm(Y ~ brain_z + X, data = d)
summary(Reg2DirectEffect)

# Using PROCESS macro 

process(d, y="Y", x="X", m= c("brain_z"), model = 4,  effsize = 1,total = 1,  normal = 1, boot = 5000, seed = 9999)  

```


### abstinence -> impulsivity  -> craving (brain only)

```{R}
# define X, Y, and M
d$X = d$abstinence_z
d$Y = d$craving_z

#Computing linear regression for PATH A
Reg2PathA <- lm(BIS_z ~ X, data = d)
summary(Reg2PathA)

#Conducting linear regression for PATH B and C'
Reg2DirectEffect <- lm(Y ~ BIS_z + X, data = d)
summary(Reg2DirectEffect)

# Using PROCESS macro 

process(d, y="Y", x="X", m= c("BIS_z"), model = 4,  effsize = 1,total = 1,  normal = 1, boot = 5000, seed = 9999)  

```